---
title: "Definitions"
output:
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    number_sections: false
    toc_depth: 1
    #code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message=FALSE,warning=FALSE, cache=TRUE)
```

*Assignment #2 - Functions*

#help()

*the help function outputs technical information about an R-topic and describes how to use such topic
```{r}
  help(list)
```

#?topic

*same as the help function
```{r}
  ?list
```

#ls()

*outputs a vector of the names of all current objects created in the global enviornment
```{r}
  ls()
```

#dir()

*outputs a vector of the names of all R-related files inside project folder
```{r}
  dir_array <-dir()
  dir_array
```

#list.files()

*output is the same as dir()
```{r}
  list.files()
```

#save()
*saves an R-object from the workspace to a file specified by user for R to understand
```{r}
  array1 <- c(1,2,3)
  save(array1, file = "abc.rda")
```

#load()

*loads a previously saved R-object file into the workspace
```{r}
  load("abc.rda")
```

#data()

*displays all available data sets preloaded into R-studio as a list object
```{r}
  data_array <- data()
  data_array[[3]][97,]
```

#library()

*displays all available R-packages to add functionality as a list object
```{r}
  library_array <- library()
  library_array[[2]][2,]
```

