---
title: "Definitions"
output:
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    number_sections: false
    toc_depth: 1
    #code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message=FALSE,warning=FALSE, cache=TRUE)
```

*Assignment #2 - Functions*

#help()

*the help function outputs technical information about an R-topic and describes how to use such topic
```{r}
  help(list)
```

#?topic

*same as the help function
```{r}
  ?list
```

#ls()

*outputs a vector of the names of all current objects created in the global enviornment
```{r}
  ls()
```

#dir()

*outputs a vector of the names of all R-related files inside project folder
```{r}
  dir_array <-dir()
  dir_array
```

#list.files()

*output is the same as dir()
```{r}
  list.files()
```

#save()
*saves an R-object from the workspace to a file specified by user for R to understand
```{r}
  array1 <- c(1,2,3)
  save(array1, file = "abc.rda")
```

#load()

*loads a previously saved R-object file into the workspace
```{r}
  load("abc.rda")
```

#data()

*displays all available data sets preloaded into R-studio as a list object
```{r}
  data_array <- data()
  data_array[[3]][97,]
```

#library()

*displays all available R-packages to add functionality as a list object
```{r}
  library_array <- library()
  library_array[[2]][2,]
```

#read.table()

*opens the file abc.txt in the working directory and saves columns as vectors using
a default column name "V"
```{r}
  table1 <- read.table("abc.txt")
  table1
```

#read.csv()

*opens the file abc.csv in the working directory and saves columns as vectors, preserving
column names defined by user in csv file.
```{r}
  table2 <- read.csv("abc.csv")
  table2$Fruits
```

#scan()

*can open a file filled with real numbers or receive input from keyboard if file equals "". Output is a vector of real numbers
```{r}
  num_scan <- scan("abc2.txt")
  num_scan
  num_scan[3]
```

#print()

*outputs a previously defined object, a single string, or the first number of a vector of numbers. Coerces all items to strings if characters are found.
```{r}
  sample_num <- c(1,2,3,4,5)
  print(sample_num)
  print(table2$Fruits)
  print("hello world")
  print(1,2,3)
```

#cat()

*similar to print() and c() by outputing all items to a single line, whether items are all numbers, all strings, or combination. Assinging cat() to an object results in a NULL set.
```{r}
  print(c(1,2,3,"hello world"))
  sample_cat <- cat(1,2,3, "hello world")
  sample_cat
```

#write.table()

*outputs a vector object to a file designated by user
```{r}
  table3 <- write.table(c(1,2,3,4,5,6), "abc3.txt")
  table3 <- read.table("abc3.txt")
  table3
```

#c()

*Combines items into a vector. Coerces all items to strings if characters are found.
```{r}
  array_samp <- c(1,2,3,4,5,6)
  array_samp
```

#from:to

*Outputs the sequence of numbers from one number to another number and increments by 1. Numbers can be in ascending or descending order, and negative
```{r}
  1:3
  3:1
  -1:-3
  1.3:1.5
```

#seq()

*Similar to from:to, with the option of incrementing or decremeting by numbers other than 1. Function is sign-sensitive.
```{r}
  seq(1,3)
  seq(10,1,-2)
```

#rep()

*Outputs a single number, string, or vector with no other arguments. If second argument is a number, item repeats for x number of times.
```{r}
  rep(1)
  rep(1,3)
  rep("hello world", 4)
```

#data.frame()

*Outputs a matrix of values depending on the format of the input. Ex: If we simply add numbers seperated by comma, the matrix will output one row of x columns. Another is to assign sets of numbers to tags which forces the sets to rows.
```{r}
  data.frame(1,2,3,4,5,6)
  data.frame(x = c(1,2,3), y = c(3,6,9), z = c(10,20,30))
```

