---
title: "R-Basics"
output:
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    number_sections: false
    toc_depth: 1
    #code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message=FALSE,warning=FALSE, cache=TRUE)
```


# Problem 1


```{r}
1+1
5/7
```

# Problem 2

helpful tip - "r eval=FALSE" in "{r}"(will not run code but display it)

```{r}
rich <- 1
rich

```

# Problem 3


```{r}
array1 <- 1
for(i in 2:100){
  array1 <-c(array1, i)
}
array1
```
*returns a set of numbers (all real numbers)

**alternatively**
```{r}
array2 <- seq(1,100)
array2
```
*returns a set of integers

# Problem 4

```{r}
array3 <- sum(seq(1,100))
array3
```

# Problem 5 

```{r}
mySum<- function (x,y) {
  sumVector <- sum(seq(x,y))
  return(sumVector)
}

mySum(1,5)
```

#Problem 6

```{r}
array4 <- seq(1,100,2)
array4
```

**alternatively**

for loop can iterate differently (1:100) or seq(2,100,2) or using array of strings
```{r}
array5 <- c()

for(i in 1:100){
  if(i%%2 > 0){
    array5<-c(array5,i)
  }
}
array5

```

#Problem 7

Use a vector and take advantage of NA and NULL to remove unwanted indices (non prime numbers)
```{r}
#making a sequence of numbers
vec_num <- c(2:1000)

#search the vector to tag numbers divisible by 2, 3, 5, and 7
j <- 1
while(j <= 998){
  vec_num[2+j] <- NA
  j <- j+2
}

k <- 2
while(k <= 997){
  vec_num[3+k] <- NA
  k <- k+3
}

m <- 4
while(m <= 995){
  vec_num[5+m] <- NA
  m <- m+5
}

n <- 6
while(n <= 993){
  vec_num[7+n] <- NA
  n <- n+7
}

# Create new vector without tagged numbers

newvec_num <- c()

for(a in 1:length(vec_num)){
  if(!is.na(vec_num[a])){
    newvec_num <- c(newvec_num, vec_num[a])
  }
  
}

newvec_num

```

# Problem 8

```{r}
  u <- runif(100)
  u
```

# Problem 9

```{r}
  alpha <- runif(100, min = 1, max = 9)
  alpha
```

# Problem 10

```{r}
  aa <- c(10,20,30,40,50,50)
  mean_of_X <- function(x) {
    return (sum(x)/length(x))
  }
  
  mean_of_X(aa)
  
#  mode_of_X <- function(x) {
#    for(i in 1:length(x)) {
#      for(j in 1:length(x)) {
#        if(x[i])
#      }
#    }
#  }

```

