---
title: "R-Basics"
output:
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    number_sections: false
    toc_depth: 1
    #code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message=FALSE,warning=FALSE, cache=TRUE)
```


# Problem 1


```{r}
1+1
5/7
```

# Problem 2

helpful tip - "r eval=FALSE" in "{r}"(will not run code but display it)

```{r}
rich <- 1
rich

```

# Problem 3


```{r}
array1 <- 1
for(i in 2:100){
  array1 <-c(array1, i)
}
array1
```
*returns a set of numbers (all real numbers)

**alternatively**
```{r}
array2 <- seq(1,100)
array2
```
*returns a set of integers

# Problem 4

```{r}
array3 <- sum(seq(1,100))
array3
```

# Problem 5 

```{r}
mySum<- function (x,y) {
  sumVector <- sum(seq(x,y))
  return(sumVector)
}

mySum(1,5)
```

#Problem 6

```{r}
array4 <- seq(1,100,2)
array4
```

**alternatively**

for loop can iterate differently (1:100) or seq(2,100,2) or using array of strings
```{r}
array5 <- c()

for(i in 1:100){
  if(i%%2 > 0){
    array5<-c(array5,i)
  }
}
array5

```

#Problem 7

Use a vector and take advantage of NA and NULL to remove unwanted indices (non prime numbers)
```{r}
#making a sequence of numbers
vec_num <- c(2:1000)

#search the vector to tag numbers divisible by 2, 3, 5, and 7
j <- 1
while(j <= 998){
  vec_num[2+j] <- NA
  j <- j+2
}

k <- 2
while(k <= 997){
  vec_num[3+k] <- NA
  k <- k+3
}

m <- 4
while(m <= 995){
  vec_num[5+m] <- NA
  m <- m+5
}

n <- 6
while(n <= 993){
  vec_num[7+n] <- NA
  n <- n+7
}

# Create new vector without tagged numbers

newvec_num <- c()

for(a in 1:length(vec_num)){
  if(!is.na(vec_num[a])){
    newvec_num <- c(newvec_num, vec_num[a])
  }
  
}

newvec_num

```

# Problem 8

```{r}
  u <- runif(100)
  u
```

# Problem 9

```{r}
  alpha <- runif(100, min = 1, max = 9)
  alpha
```

# Problem 10

```{r}
  aa <- c(10,20,20,30,40,50,60)

  mean_of_X <- function(x) {
    return (sum(x)/length(x))
  }
  
  mean_of_X(aa)
  
  mode_of_X <- function(x){
    temp_high_num <- c()
    count <- 0
    for(i in 1:length(x))   {
      for(j in 1:length(x))    {
        if(x[i] == x[j]) {
            count = count + 1
          }
      }
      if(count > 1) {
        temp_high_num <- x[i]
      }
      count <- 0
    }
      return(temp_high_num)
  }
  
  mode_of_X(aa)
  
  median_of_X <- function(x){
    if(length(x)%%2 > 0) {
      med <- x[((length(x)-1)/2) + 1]
    }
      else{
        med <- (x[(length(x)/2)] + x[(length(x)/2) + 1])/2
      }
    
    return(med)
  }
  
  median_of_X(aa)
  
  range_of_X <- function(x){
    rge <- x[length(x)] - x[1]
    return(rge)
  }
  
  range_of_X(aa)
  
  std_dev_of_X <- function(x) {
    avg <- mean_of_X(x)
    sumOfsquares <- 0
    
    for(i in 1:length(x)) {
      sumOfsquares = sumOfsquares + (x[i] - avg)^2
    }
    
    std_dev <- sqrt(sumOfsquares/(length(x)-1))
    
    return(std_dev)
  }
  
  std_dev_of_X(aa)

```

# Problem 11

```{r}
  bb <- "adrthbjyiop"
  dd <- unlist(strsplit(bb, ""))
  print(length(dd))
```

# Problem 12

```{r}
  cc <- "This is a sentence"
  ee <- unlist(strsplit(cc, split = " "))
  print(length(ee))
```

# Problem 13

```{r}
  paragraph <- "Four score and seven years ago our fathers brought forth on this    continent, a new nation, conceived in Liberty, and dedicated to the proposition that all men are created equal. Now we are engaged in a great civil war, testing whether that nation, or any nation so conceived and so dedicated, can long endure. We are met on a great battle-field of that war."

  sep_paragraph <- unlist(strsplit(paragraph, split = ".", fixed = TRUE))
  print(length(sep_paragraph))
```

# Problem 14 (alternative solution)

```{r}
 search_char <- function(x, letter){
    dd <- unlist(strsplit(x, ""))
    count <- 0
    for(i in 1:length(dd)) {
        if(dd[i] == letter){
            count = count + 1
          }

    }
      return(count)
 }

  search_char("abcdeggg", "g")
  
```

# Problem 15

```{r}
  test_word <- "hello"
  test_sentence <- "hello my name is Richard"
  split_test_sentence <- unlist(strsplit(test_sentence, split = " "))
  test_word%in%split_test_sentence
```

# Problem 16

```{r}
  current_time <- (as.numeric(Sys.time())*1000)
  print(current_time, digits = 15)
```

# Problem 17

```{r}
  current_time1 <- (as.numeric(Sys.time())*1000)
  for(i in 1:10000000){}
  current_time2 <- (as.numeric(Sys.time())*1000)
  cat(total_time = current_time2 - current_time1, "milliseconds")
```

# Problem 18

```{r}
  table2 <- read.csv("abc.csv")
  table2$Fruits
```

# Problem 19

```{r}
  table3 <- write.table(c(1,2,3,4,5,6), "abc3.txt")
  table3 <- read.table("abc3.txt")
  table3
```

# Problem 20

```{r}
  random_matrix <- matrix(runif(100, min = 1, max = 9), ncol=20,nrow=20)
  random_matrix[,19]
```

# Problem 21

```{r}
  matrixA <- data.frame(x = c(1,2,3), y = c(3,6,9), z = c(10,20,30))
  write.csv(matrixA, file = "sample_matrix.csv", row.names = FALSE)
```

# HARDER PROBLEMS

Fizzbuzz

```{r, eval=FALSE}
  vec_num <- c(1:100)

  k <- 1
  while(k <= 100) {    
    if(vec_num[k]%%3==0 || vec_num[k]%%5==0) {
      if(vec_num[k]%%3==0 && vec_num[k]%%5== 0) {
          print("FizzBuzz")
          k <- k + 1
        }
          else if(vec_num[k]%%5 == 0) {
            print("Buzz")
            k <- k + 1
          }
        else {
          print("Fizz")
          k <- k + 1
        }
    }
    
      else {
        print(vec_num[k])
        k <- k + 1          
      }

}

```

Frequency Count

```{r}
 search_char <- function(x, letter){
    dd <- unlist(strsplit(x, ""))
    count <- 0
    for(i in 1:length(dd)) {
        if(dd[i] == letter){
            count <- count + 1
          }

    }
      return(count)
 }

  phrase <- "I love R studio"
  search_phrase <- unlist(strsplit(phrase, ""))
  found_number <- c()
  
  for(i in 1:length(search_phrase)) {
    found_number <- c(found_number, search_char(phrase, search_phrase[i]))
  }
  
  xx <- data.frame(search_phrase, found_number)
  print(unique(xx))
```

Test Random Number Generator

```{r}

  rand_num <- runif(1000000,0,100)

  hist(rand_num, breaks = 101)
    
```

