---
title: "Reducing Implicit Bias"
output: 
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    number_sections: false
    toc_depth: 1
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message = FALSE)
```

# Abstract
  There is evidence by Hu (2015) on the unlearning of social bias by reinforcing a counterbias behavior during slow-wave/REM sleep - the optimal time frame to consolidate new memories. The reproduced analysis pertains to the interaction of cued and uncued reduction bias before sleep (prenap), then, one week later (delayed). A power analysis shows that the original effect size may only be observed approximately 14% of the time on average.
```{r}
#downloads csv data into R
aaa3841_OpenData_withNA <- read.csv("~/GitHub(assignments)/midterm/midterm/aaa3841_OpenData_withNA.csv")

```

# Reproducing analysis of: Unlearning implicit social biases during sleep (2015)

The experiment on unlearning social bias shows the relevance of sleep playing a role in
the dissaption of a pre-existing implicit bias, Hu et al. (2015). The rationale comes from
auditory information reinforcing learning behavior during slow wave sleep, causing a targeted
memory reactivation. As a result of the auditory cue, counterbias behavior in participants
was expected to increase. In addition, the memory reactivation should result in a sustained
reduction in social bias, in that the measured bias does not differ between prenap and delay.
The implications of such an effect may be relevant for unlearning other unwanted habits.
The original data was downloaded from https://osf.io/b3k9a/
```{r}
#selecting data for analysis
cue_prenap = aaa3841_OpenData_withNA$Cued_Prenap
cue_delay = aaa3841_OpenData_withNA$Cued_Delayed
uncue_prenap = aaa3841_OpenData_withNA$Uncued_Prenap
uncue_delay = aaa3841_OpenData_withNA$Uncued_Delayed
```

```{r}
#preparing new data table of the two factors with IAT scores for each subject
cue_factor = rep(c("C","U"), each = 80)
time_factor = rep(rep(c("D","P"), each = 40), 2)
subjects = aaa3841_OpenData_withNA$Ppt.
sub = as.factor(subjects)
subject_dv = c()
subject_dv =  c(cue_delay[1:length(cue_delay)], cue_prenap[1:length(cue_prenap)], 
                uncue_delay[1:length(uncue_delay)], uncue_prenap[1:length(uncue_prenap)])
```

```{r}
#Preparing data for analysis by filtering data for NA's(bad subjects)
all_data = data.frame(sub, subject_dv, cue_factor, time_factor)
library(dplyr)
nas <- all_data %>%
  filter(is.na(subject_dv)==TRUE)
nas$sub<-as.factor(as.numeric(nas$sub))
bad_subjects <- levels(nas$sub)

#filtered data
clean_df <- all_data %>%
              filter(sub %in% bad_subjects == FALSE)

```
# Results

IAT bias scores were grouped in R-studio for each participant in order to perform a
2x2 (cue/uncue vs prenap/delay) repeated measure ANOVA. The group means are displayed
in Figure 1 followed by the ANOVA results in Figure 2. **There was a main effect of Time,
F(1, 37) = 5.76, MSE = 0.15, p = .022.** Mean IAT scores remained lower after
training, between prenap testing (0.447) and testing one week later (0.334). There was not a
main effect of cue factor after learning, F(1, 37) = 0.34, MSE = 0.28, p = .566.
**The two-way interaction between Time and Cue after training was significant,
F(1, 37) = 4.67, MSE = 0.07, p = .037.** The difference in prenap and delay scores
was significant only when the cue was not initiated, indicating that memory reactivation
sustained after counterbias training.

**Figure 1.**
```{r}
library(xtable)
  means = c(mean(cue_prenap), mean(cue_delay, na.rm = TRUE), 
            mean(uncue_prenap), mean(uncue_delay, na.rm = TRUE))
  SEM = c(sd(cue_prenap)/38, sd(cue_delay, na.rm=TRUE)/38, 
          sd(uncue_prenap)/38, sd(uncue_delay, na.rm=TRUE)/38)
  group_means = data.frame(GroupMean=means, Cue=c("Cue","Cue","Uncue","Uncue"), 
                         Time=c("Prenap","Delay","Prenap","Delay"), SEM)
  knitr::kable(xtable(group_means))
```

```{r}
#Visualize how the average bias score changes between conditions 
group_means = data.frame(all_means=means, cue_f=c("Cue","Cue","Uncue","Uncue"), 
                         time_cond=c("Prenap","Delay","Prenap","Delay"), SEM)
library(ggplot2)
ggplot(group_means, aes(x=cue_f,
                          y=all_means, 
                          group=time_cond,
                          fill=time_cond))+
  geom_bar(stat="identity",position="dodge")+
  theme_classic(base_size=12)+
  xlab("Cue factor") +
  ylab("Mean IAT scores")+
  geom_errorbar(aes(ymin=all_means-SEM,
                    ymax=all_means+SEM),
                position=position_dodge(width=0.9),
                width=.2,
                color="black")+
  coord_cartesian(ylim=c(0,1))
```

**Figure 2.**
```{r}
# 2-way between subjects anova
library(xtable)
aov_out = aov(subject_dv~cue_factor*time_factor+Error(sub/(cue_factor*time_factor)), clean_df)
knitr::kable(xtable(summary(aov_out)))
```
# Power Analysis

The replicated interaction was due to a significant difference of the time factor in the
uncued group. Since an F-value is equal to the square of a t-value, we can simulate power by
simulating a t-test on the difference of time factor. The mean score difference between
prenap and delay is (0.143) for cue and uncue. We applied this difference to our simulation
assuming equal variance in our two samples (n = 38), and calculated an estimated power of
0.136, meaning the observed effect size may only occur 13.6% of the time.
```{r}
# Sim_power is a function to run a simulated t-test
sim_power <- function(x){
  A <- rnorm(n=38,mean=0, sd=1)
  B <- rnorm(n=38,mean=(0+x), sd=1)
  return(t.test(A,B,var.equal=TRUE, paired = TRUE)$p.value)
}

# vector of effect sizes to be simulated
effect_sizes <- seq(.1,2,.1)
# run a power simulation for each effect size 1000 times
power <- sapply(effect_sizes, 
                FUN = function(x) {
                  sims <- replicate(1000,sim_power(x))
                  sim_power <- length(sims[sims<.05])/length(sims)
                  return(sim_power)})

```

```{r}
# combine results of simulation into a data frame
plot_df <- data.frame(effect_sizes,power)
```

```{r}
# plot the power curve
ggplot(plot_df, aes(x=effect_sizes,
                    y=power))+
  geom_point()+
  geom_line()
```

# References

Hu, X., Antony, J. W., Creery, J. D., Vargas, I. M., Bodenhausen, G. V., & Paller, K. A.
(2015). Unlearning implicit social biases during sleep, 348 (6238), 1013–1015.
doi:10.1126/science.aaa3841
